### 2024-10-08 manuscript/ms.tex
Working on methods and results.

%&<latex>
\documentclass[letterpaper,12pt]{article}
\pdfoutput=1

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% preamble %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{../utils/preamble.tex}
\input{../utils/environments.tex}
\input{../utils/macros.tex}
\input{../utils/notation.tex}

\newcommand{\ifdoublespacing}[2]{#1}
% \newcommand{\ifdoublespacing}[2]{#2}
\newcommand{\iflinenumbers}[2]{#1}
% \newcommand{\iflinenumbers}[2]{#2}
% \newcommand{\ifragged}[2]{#1}
\newcommand{\ifragged}[2]{#2}

\captionsetup[figure]{listformat=figList}

\usepackage[round]{natbib}
\usepackage{amsmath}
\title{The great codivergence bake off}

\author[1]{Tanner Myers}
\author[1]{Randy Klabacka}
\author[1]{Hailey Whitaker}
\author[1]{Tashitso Anamza}
\author[1]{Claire Tracy}
\author[1]{Jamie R.\ Oaks \thanks{Corresponding author: \href{mailto:joaks@auburn.edu}{\tt joaks@auburn.edu}}}
\affil[1]{Department of Biological Sciences \& Museum of Natural History, Auburn University, Auburn, Alabama 36849}

\date{\today}

\makeatletter
\let\msTitle\@title
\let\msAuthor\@author
\let\msDate\@date
\makeatother
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\ifragged{
\raggedright
}{}

\iflinenumbers{
\begin{linenumbers}
}{}

\textbf{Running head}: \uppercase{Codiv bake off}

{\let\newpage\relax\maketitle}

\newpage

\ifdoublespacing{
\doublespacing
}{}

\begin{abstract}
    Abstract here \ldots

    \vspace{6pt}
    \noindent\textbf{KEY WORDS:}
\end{abstract}

\newpage

\section{Introduction}

\section{Methods}

\subsection{Analyses of simulated data}

We simulated loci for 5 diploid individuals (i.e., 10 gene
copies) per population from 5 pairs of diverging populations.
Each pair shared a common ancestral population with a constant effective size
\epopsize[\rootpopindex]
that diverged at time
\comparisonetime{}
into two descendant populations with constant effective sizes
\epopsize[\descendantpopindex{1}]
and 
\epopsize[\descendantpopindex{2}].
Each locus we simulated comprised 500 linked characters with no intralocus
recombination.
We simulated datasets comprising 1, 2, 5, and 20 independent loci.
For each number of loci, we simulated datasets under three divergence
scenarios:
\begin{enumerate}
    \item All five pairs of populations diverged independently.
    \item All five pairs of populations diverged simultaneously.
    \item The number of divergence times (\nevents) and the assignment of the
        five pairs of populations to the divergene times (\divtimesets) vary
        according to a Dirichlet-process model with a concentration parameter
        of 2.22543.
\end{enumerate}
For each of the 12 simulation conditions (4 numbers of loci $\times$ 3
divergence scenarios),
we simulated 1000 datasets.
Each unique divergence time was drawn from an exponential distribution,
$\comparisonetime \sim \dexponential{0.01}$.
We constrained the mutation rate (\murate{}) to 1 across the five pairs of
populations, such that divergence times and effective population sizes were
scaled by the mutation rate.

\subsubsection{Full-likelihood analyses}

To simulate datasets and analyze them under a full-likelihood Bayesian model of
population divergences, we used the \simcoevolity and \ecoevolity tools of the
\ecoevolity software package
\citep[Version 0.3.2, Commit c128046;][]{Oaks2018ecoevolity,Oaks2018paic}.

The mutation-scaled effective sizes of the descendant populations
($\epopsize\murate$) were drawn from a gamma distribution with a shape of 10
and mean of 0.0005,
$\epopsize{}\murate{} \sim \dgamma{10}{0.0005}$.
The effective size of the ancestral population relative to the mean of
its descendants was drawn from a gamma distribution with a shape of 40 and mean of 1,
$\rootrelativepopsize \sim \dgamma{40}{1.0}$.

When analyzing each simulated \dataset with \ecoevolity, we ran four MCMC
chains for 30,000 generations sampling every 20 generations.
We used the last 1000 samples from each of the four chains to approximate the
posterior distribution of the model given each simulated dataset.
To assess convergence and mixing of the MCMC chains, we calculated the
potential scale reduction factor \citep[PSRF; the square root of Equation 1.1
in][]{Brooks1998} and effective sample size \citep[ESS;][]{Gong2014} for all
continuous parameters and the log likelihood using
\pycoevolity
\citep[Version 0.2.11 Commit 85ea44b;][]{Oaks2018ecoevolity,PycoevolityRepoOnline}.
% When plotting results, we highlight any simulation replicates that have a
% $\textrm{PSRF} > 1.2$.

\subsubsection{ABC analyses}

To simulate datasets and analyze them under a similar model estimated via
approximate-Bayesian computation (ABC),
we used \pymsbayes
\citep[Version 0.3.7, Commit 54f720df;][]{Oaks2014dpp},
which is a Python multiprocessing wrapper around dpp-msbayes
\citep{Oaks2014dpp}, which is a modification of
MTML-msBayes \citep{Huang2011}.
In this software, the distribution for effective population sizes is placed
on the parameter $4\epopsize{}\murate{}$, for which we used
a gamma distribution with a shape of 10 and a mean of 0.002.
Note, this is equivalent to the distribution we used for $\epopsize{}\murate{}$
in \ecoevolity.

\subsection{Empirical analyses}

\section{Results}

\subsection{Simulation results}

The potential scale reduction facter (PSRF) was less than X and the effective
sample size (ESS) was greater than Y for all parameters and the log likelihood
across all simulations
(Fig.\ S\labelcref{fig:psrfdivtime,fig:psrfrootpopsize,fig:psrflnl,fig:essdivtime,fig:essrootpopsize,fig:esslnl}).

\begin{figure}[htbp]
    \begin{center}
        \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{../images/from-project-repo/plots/tex-plot-grids/grid-model-distance-cropped.pdf}
        \caption{
            \Ecoevolity estimates the partitioning of population pairs to
            divergence times (the divergence model) more accurately and
            precisely than \dppmsbayes across all simulated numbers of loci
            (columns) and whether the divergence model was drawn from a
            Dirichlet process (top row), or constrained so that all pairs of
            populations diverged independently (middle row) or simultaneously
            (bottom row).
            Divergence model estimation error was quantified for each
            simulation replicate using the mean partition distance
            \citep{Regnier1983,Gusfield2002} between the true divergence model
            and the posterior samples of divergence models.
            Partition distances was calculated using the Hungarian (or
            Kuhn-Munkres) algorithm
            \citep{Kuhn1955,Munkres1957}
            as implemented in \pycoevolity
            \citep[Version 0.2.11 Commit 85ea44b;][]{Oaks2018ecoevolity,PycoevolityRepoOnline},
            which relies on the Munkres Python package
            \citep[Version 1.1.4;][]{Clapper2020}.
            For each violin plot, the sample size (number of simulation
            replicates) is shown beneath and the mean is represented by the
            white dot and shown above.
            The box shows the upper and lower quartiles and the brackets
            show the 2.5th and 97.5th percentiles.
            At the top center of each plot, the results of a Mann-Whitney U
            test \citep{MannWhitney1947} are shown comparing mean model errors
            between \ecoevolity and \dppmsbayes;
            the probability of superiority \citep[PoS, the probability that
            \ecoevolity has a lower mean model error than \dppmsbayes for a
            random simulation replicate drawn from
            each;][]{WolfeHogg1971,Grissom1994} and p-value is provided.
        }
        \label{fig:modelerror}
    \end{center}
\end{figure}

\begin{figure}[htbp]
    \begin{center}
        \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{../images/from-project-repo/plots/tex-plot-grids/grid-div-time-error-cropped.pdf}
        \caption{
            \Ecoevolity better estimates divergence times than \dppmsbayes
            across all simulated numbers of loci (columns) and whether the
            divergence model was drawn from a Dirichlet process (top row), or
            constrained so that all pairs of populations diverged independently
            (middle row) or simultaneously (bottom row).
            Divergence time error was quantified for each simulation replicate
            as the sum of the difference between the true and posterior mean
            divergence time across the five pairs of populations.
            For each violin plot, the sample size (number of simulation
            replicates) is shown beneath and the mean is represented by the
            white dot and shown above.
            The box shows the upper and lower quartiles and the brackets
            show the 2.5th and 97.5th percentiles.
            At the top center of each plot, the results of a Mann-Whitney U
            test \citep{MannWhitney1947} are shown comparing mean model errors
            between \ecoevolity and \dppmsbayes;
            the probability of superiority \citep[PoS, the probability that
            \ecoevolity has a lower mean model error than \dppmsbayes for a
            random simulation replicate drawn from
            each;][]{WolfeHogg1971,Grissom1994} and p-value is provided.
        }
        \label{fig:divtimeerror}
    \end{center}
\end{figure}

\begin{figure}[htbp]
    \begin{center}
        \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{../images/from-project-repo/plots/free-div-time-scatter-cropped.pdf}
        \caption{
            \Ecoevolity (top row) estimates divergence times more accurately
            and precisely than \dppmsbayes (bottom row) across all simulated
            numbers of loci (columns) for five pairs of populations simulated
            to have diverged according to a Dirichlet process.
            Each plotted circle and associated error bars represent the posterior mean
            and 95\% credible interval.
            \accuracyscatterplotannotations{\comparisonetime{}}
        }
        \label{fig:divtimeerrordp}
    \end{center}
\end{figure}

\subsection{Empirical results}

\section{Discussion}

\section{Acknowledgments}

This work was supported by funding provided to JRO from the National Science
Foundation (NSF grant number DEB 1656004).
The computational work was made possible by the Auburn University (AU) Hopper
Cluster supported by the AU Office of Information Technology and a grant of
high-performance computing resources and technical support from the Alabama
Supercomputer Authority.
This paper is contribution number \highLight{XXX} of the Auburn University
Museum of Natural History.

%TC:ignore

\bibliographystyle{evolution}
\bibliography{references}

\iflinenumbers{
\end{linenumbers}
}{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% SUPPORTING INFO %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\setcounter{figure}{0}
\setcounter{table}{0}
\setcounter{page}{1}
\setcounter{section}{0}

\captionsetup[figure]{name=Figure S, labelformat=noSpace, listformat=sFigList}
\captionsetup[table]{name=Table S, labelformat=noSpace}

\singlespacing

\section*{Supporting Information}
\hangindent=1cm
\noindent Title: \msTitle

\bigskip
{\noindent Authors: \msAuthor}

\newpage
\singlespacing

\input{si-body.tex}
\clearpage

\input{si-tables.tex}
\clearpage

\input{si-figures.tex}

%TC:endignore

\end{document}