### 2024-11-20 manuscript/ms.tex
Mostly working on empirical methods and results.
Also making edits to simulation methods and results.

%&<latex>
\documentclass[letterpaper,12pt]{article}
\pdfoutput=1

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% preamble %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{../utils/preamble.tex}
\input{../utils/environments.tex}
\input{../utils/macros.tex}
\input{../utils/notation.tex}

\newcommand{\ifdoublespacing}[2]{#1}
% \newcommand{\ifdoublespacing}[2]{#2}
\newcommand{\iflinenumbers}[2]{#1}
% \newcommand{\iflinenumbers}[2]{#2}
% \newcommand{\ifragged}[2]{#1}
\newcommand{\ifragged}[2]{#2}

\captionsetup[figure]{listformat=figList}

\usepackage[round]{natbib}
\usepackage{amsmath}
\title{The great codivergence bake off}

\author[1]{Tanner Myers}
\author[1]{Randy Klabacka}
\author[1]{Hailey Whitaker}
\author[1]{Tashitso Anamza}
\author[1]{Claire Tracy}
\author[1]{Jamie R.\ Oaks \thanks{Corresponding author: \href{mailto:joaks@auburn.edu}{\tt joaks@auburn.edu}}}
\affil[1]{Department of Biological Sciences \& Museum of Natural History, Auburn University, Auburn, Alabama 36849}

\date{\today}

\makeatletter
\let\msTitle\@title
\let\msAuthor\@author
\let\msDate\@date
\makeatother
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\ifragged{
\raggedright
}{}

\iflinenumbers{
\begin{linenumbers}
}{}

\textbf{Running head}: \uppercase{Codiv bake off}

{\let\newpage\relax\maketitle}

\newpage

\ifdoublespacing{
\doublespacing
}{}

\begin{abstract}
    Abstract here \ldots

    \vspace{6pt}
    \noindent\textbf{KEY WORDS:}
\end{abstract}

\newpage

\section{Introduction}

\section{Methods}

\subsection{Simulating data}

We simulated loci for five diploid individuals (i.e., 10 gene
copies) per population from five pairs of diverging populations.
Each pair shared a common ancestral population with a constant effective size
\epopsize[\rootpopindex]
that diverged at time
\comparisonetime{}
into two descendant populations with constant effective sizes
\epopsize[\descendantpopindex{1}]
and 
\epopsize[\descendantpopindex{2}].
Each locus we simulated comprised 500 linked characters with no intralocus
recombination.
We simulated datasets comprising 1, 2, 5, and 20 independent loci.
For each number of loci, we simulated datasets under three divergence
scenarios:
\begin{enumerate}
    \item All five pairs of populations diverged independently.
    \item All five pairs of populations diverged simultaneously.
    \item The number of divergence times (\nevents) and the assignment of the
        five pairs of populations to the divergence times (\etimesets) vary
        according to a Dirichlet-process model with a concentration parameter
        of 2.22543, which corresponds to a mean of 3 divergence times.
\end{enumerate}
For each of the 12 simulation conditions (4 numbers of loci $\times$ 3
divergence scenarios),
we simulated 1000 datasets.
Each unique divergence time was drawn from an exponential distribution,
$\comparisonetime \sim \dexponential{0.01}$.
We constrained the mutation rate (\murate{}) to 1 across the five pairs of
populations, such that divergence times and effective population sizes were
scaled by the mutation rate.

To simulate the datasets for analysis under a full-likelihood Bayesian model,
we used the \simcoevolity tool of the \ecoevolity software package.
\citep[Version 0.3.2, Commit c128046;][]{Oaks2018ecoevolity,Oaks2018paic}.
The mutation-scaled effective sizes of the descendant populations
($\epopsize\murate$) were drawn from a gamma distribution with a shape of 10
and mean of 0.0005,
$\epopsize{}\murate{} \sim \dgamma{10}{0.0005}$.
The effective size of the ancestral population relative to the mean of
its descendants was drawn from a gamma distribution with a shape of 40 and mean of 1,
$\rootrelativepopsize \sim \dgamma{40}{1.0}$.

To simulate datasets for analysis with
approximate-Bayesian computation (ABC),
we used \pymsbayes
\citep[Version 0.3.7, Commit 54f720df;][]{Oaks2014dpp},
which is a Python multiprocessing wrapper around \dppmsbayes
\citep{Oaks2014dpp}, which is a modification of
MTML-msBayes \citep{Huang2011}.
In this software, the distribution for effective population sizes is placed
on the parameter $4\epopsize{}\murate{}$, for which we used
a gamma distribution with a shape of 10 and a mean of 0.002.
This is equivalent to the distribution we used for $\epopsize{}\murate{}$ with
\ecoevolity.
When summarizing and plotting effective population size for \pymsbayes, we
divided all true and estimated values by four for direct comparability with
\ecoevolity.

\subsection{Full-likelihood analyses of simulated data}

We analyzed each dataset simulated by \simcoevolity with \ecoevolity
\citep[Version 0.3.2, Commit c128046;][]{Oaks2018ecoevolity,Oaks2018paic}
under a Dirichlet-process model with a concentration parameter of 2.22543.
For prior distributions on divergence times and effective population sizes, we
used the same distributions described above for the simulations.
We ran four MCMC chains for 30,000 generations sampling every 20 generations.
To approximate the posterior distribution of the model given each simulated
dataset, we used the last 1000 samples from each of the four chains.
To assess convergence and mixing of the MCMC chains, we calculated the
potential scale reduction factor \citep[PSRF; the square root of Equation 1.1
in][]{Brooks1998} and effective sample size \citep[ESS;][]{Gong2014} for all
continuous parameters and the log likelihood using
\pycoevolity
\citep[Version 0.2.11 Commit 85ea44b;][]{Oaks2018ecoevolity,PycoevolityRepoOnline}.

\subsubsection{ABC analyses of simulated data}

To analyze each dataset simulated by \pymsbayes, we used the ABC rejection
algorithm implemented in \dppmsbayes via \pymsbayes \citep[Version 0.3.7,
Commit 54f720df;][]{Oaks2014dpp},
under a Dirichlet-process model with a concentration parameter of 2.22543, and
prior distributions on divergence times and effective population sizes that
match the simulation distributions described above.
We simulated 1 million samples to approximate the joint prior distribution of
the model.
To approximate the posterior given each simulated dataset, we retained the 2000
prior samples with the smallest Euclidean distance for the following four
summary statistics standardized using the mean and standard deviation from the
prior sample
\citep{Oaks2014dpp}:
$\pi$ \cite{Tajima1983}, $\theta_W$ \cite{Watterson1975}, $\pi_{net}$
\cite{Takahata1985}, and $SD(\pi-\theta_W)$ \cite{Tajima1989}.

\subsection{Re-analyses of biological data}

To assess the effect of a using a full-likelihood approach, we reanalyzed two
datasets with previously published results of using the ABC approach to
estimating shared divergence times.

In their study into the factors affecting the divergence of flightless darkling
beetles across the Aegean Archipelago,
Papadopoulou and Knowles \citeyear{PapadopoulouKnowles2015} used \dppmsbayes to
estimate the relative divergence times of 13 pairs of populations.
\jrocomment{We should probably describe the dataset a bit; number of loci, etc.}

To test the prediction of clustered divergences driven by Pleistocene glacial
cycles Oaks et al.\ \citeyear{Oaks2012} used MTML-msBayes \cite{Huang2011} to
estimate the divergence times of 22 pairs of populations from across the
Philippine Islands.
Oaks \citeyear{Oaks2014dpp} reanalyzed these same data using \dppmsbayes.
\jrocomment{We should probably describe the dataset a bit; number of loci, etc.}

\jrocomment{Jamie needs to add analysis details here}.

\section{Results}

\subsection{Simulation results}

We found good convergence and mixing of the MCMC chains across the \ecoevolity
analyses of the 12,000 simulated dataset.
The potential scale reduction factor (PSRF) was less than 1.01 and the
effective sample size (ESS) was greater than 600 for all parameters and the log
likelihood across all simulations
(\figs S\labelcref{fig:psrfdivtime,fig:psrfrootpopsize,fig:psrflnl,fig:essdivtime,fig:essrootpopsize,fig:esslnl}).

Across all four numbers of loci and three divergence scenarios, \ecoevolity
performed better at estimating the divergence model (\ie, the number of
divergence times and the assignment of the five pairs of populations to those
divergences) than \dppmsbayes, and the differences were significant according
to a Mann-Whitney U test \citep[\fig~\ref{fig:modelerror};][]{MannWhitney1947}.
The better performance of \ecoevolity is small when there is only one 500-base
locus, but the performance difference increases as the number of loci increase.
This is largely due to the fact that the ABC method (\dppmsbayes) does not gain
much information from additional loci, whereas the full-likelihood method
(\ecoevolity) does (\fig~\ref{fig:modelerror}).
The same pattern is true when looking at estimation error of divergence times
(\figs \ref{fig:divtimeerror}, \ref{fig:divtimescatterdp}, \& S\ref{fig:divtimescatter}),
the number of divergence times
(\fig S\ref{fig:nevents}),
and the effective population sizes 
(\figs S\labelcref{fig:ancpopsizeerror,fig:leafpopsizescatter,fig:rootpopsizescatter}).

\begin{figure}[htbp]
    \begin{center}
        \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{../images/from-project-repo/plots/tex-plot-grids/grid-model-distance-cropped.pdf}
        \caption{
            \Ecoevolity estimates the partitioning of population pairs to
            divergence times (the divergence model) more accurately and
            precisely than \dppmsbayes across all simulated numbers of loci
            (columns) and whether the divergence model was drawn from a
            Dirichlet process (top row), or constrained so that all pairs of
            populations diverged independently (middle row) or simultaneously
            (bottom row).
            Divergence model estimation error was quantified for each
            simulation replicate using the mean partition distance
            \citep{Regnier1983,Gusfield2002} between the true divergence model
            and the posterior samples of divergence models.
            Partition distances was calculated using the Hungarian (or
            Kuhn-Munkres) algorithm
            \citep{Kuhn1955,Munkres1957}
            as implemented in \pycoevolity
            \citep[Version 0.2.11 Commit 85ea44b;][]{Oaks2018ecoevolity,PycoevolityRepoOnline},
            which relies on the Munkres Python package
            \citep[Version 1.1.4;][]{Clapper2020}.
            For each violin plot, the sample size (number of simulation
            replicates) is shown beneath and the mean is represented by the
            white dot and shown above.
            The box shows the upper and lower quartiles and the brackets
            show the 2.5th and 97.5th percentiles.
            At the top center of each plot, the results of a Mann-Whitney U
            test \citep{MannWhitney1947} are shown comparing mean model errors
            between \ecoevolity and \dppmsbayes;
            the probability of superiority \citep[PoS, the probability that
            \ecoevolity has a lower mean model error than \dppmsbayes for a
            random simulation replicate drawn from
            each;][]{WolfeHogg1971,Grissom1994} and p-value is provided.
        }
        \label{fig:modelerror}
    \end{center}
\end{figure}

\begin{figure}[htbp]
    \begin{center}
        \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{../images/from-project-repo/plots/tex-plot-grids/grid-div-time-error-cropped.pdf}
        \caption{
            \Ecoevolity better estimates divergence times than \dppmsbayes
            across all simulated numbers of loci (columns) and whether the
            divergence model was drawn from a Dirichlet process (top row), or
            constrained so that all pairs of populations diverged independently
            (middle row) or simultaneously (bottom row).
            Divergence time error was quantified for each simulation replicate
            as the sum of the difference between the true and posterior mean
            divergence time across the five pairs of populations.
            For each violin plot, the sample size (number of simulation
            replicates) is shown beneath and the mean is represented by the
            white dot and shown above.
            The box shows the upper and lower quartiles and the brackets
            show the 2.5th and 97.5th percentiles.
            At the top center of each plot, the results of a Mann-Whitney U
            test \citep{MannWhitney1947} are shown comparing mean model errors
            between \ecoevolity and \dppmsbayes;
            the probability of superiority \citep[PoS, the probability that
            \ecoevolity has a lower mean model error than \dppmsbayes for a
            random simulation replicate drawn from
            each;][]{WolfeHogg1971,Grissom1994} and p-value is provided.
        }
        \label{fig:divtimeerror}
    \end{center}
\end{figure}

\begin{figure}[htbp]
    \begin{center}
        \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{../images/from-project-repo/plots/free-div-time-scatter-cropped.pdf}
        \caption{
            \Ecoevolity (top row) estimates divergence times more accurately
            and precisely than \dppmsbayes (bottom row) across all simulated
            numbers of loci (columns) for five pairs of populations simulated
            to have diverged according to a Dirichlet process.
            Each plotted circle and associated error bars represent the posterior mean
            and 95\% credible interval.
            \accuracyscatterplotannotations{\comparisonetime{}}
        }
        \label{fig:divtimescatterdp}
    \end{center}
\end{figure}

\subsection{Empirical results}

\jrocomment{For now I am just dropping plots below.}

\begin{figure}[htbp]
    \begin{center}
        \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{../images/from-project-repo/empirical-plots/mediterranean-all-pairs/pyco-sumevents-mediterranean-all-pairs-pycoevolity-nevents-cropped.pdf}
        \caption{
            \Ecoevolity results for all Mediterranean pairs.
        }
        \label{fig:neventsmediall}
    \end{center}
\end{figure}

\begin{figure}[htbp]
    \begin{center}
        \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{../images/from-project-repo/empirical-plots/mediterranean-soil-pairs/pyco-sumevents-mediterranean-soil-pairs-pycoevolity-nevents-cropped.pdf}
        \caption{
            \Ecoevolity results for soil Mediterranean pairs.
        }
        \label{fig:neventsmedisoil}
    \end{center}
\end{figure}

\begin{figure}[htbp]
    \begin{center}
        \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{../images/from-project-repo/empirical-plots/philippines-all-pairs/pyco-sumevents-philippines-all-pairs-pycoevolity-nevents-cropped.pdf}
        \caption{
            \Ecoevolity results for all Philippines pairs.
        }
        \label{fig:neventsphilall}
    \end{center}
\end{figure}

\begin{figure}[htbp]
    \begin{center}
        \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{../images/from-project-repo/empirical-plots/philippines-negros-panay-pairs/pyco-sumevents-philippines-negros-panay-pairs-pycoevolity-nevents-cropped.pdf}
        \caption{
            \Ecoevolity results for Negros-Panay pairs.
        }
        \label{fig:neventsphilnp}
    \end{center}
\end{figure}

\section{Discussion}

\section{Acknowledgments}

This work was supported by funding provided to JRO from the National Science
Foundation (NSF grant number DEB 1656004).
The computational work was made possible by the Auburn University (AU) Hopper
Cluster supported by the AU Office of Information Technology and a grant of
high-performance computing resources and technical support from the Alabama
Supercomputer Authority.
This paper is contribution number \highLight{XXX} of the Auburn University
Museum of Natural History.

%TC:ignore

\bibliographystyle{evolution}
\bibliography{references}

\iflinenumbers{
\end{linenumbers}
}{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% SUPPORTING INFO %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\setcounter{figure}{0}
\setcounter{table}{0}
\setcounter{page}{1}
\setcounter{section}{0}

\captionsetup[figure]{name=Figure S, labelformat=noSpace, listformat=sFigList}
\captionsetup[table]{name=Table S, labelformat=noSpace}

\singlespacing

\section*{Supporting Information}
\hangindent=1cm
\noindent Title: \msTitle

\bigskip
{\noindent Authors: \msAuthor}

\newpage
\singlespacing

\input{si-body.tex}
\clearpage

\input{si-tables.tex}
\clearpage

\input{si-figures.tex}

%TC:endignore

\end{document}