### 2025-11-26 manuscript/ms.tex
Working on manuscript.
Sketching out an introduction and adding some text about
the nature of the simulations.

%&<latex>
\documentclass[letterpaper,12pt]{article}
\pdfoutput=1

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% preamble %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{../utils/preamble.tex}
\input{../utils/environments.tex}
\input{../utils/macros.tex}
\input{../utils/notation.tex}

\newcommand{\ifdoublespacing}[2]{#1}
% \newcommand{\ifdoublespacing}[2]{#2}
\newcommand{\iflinenumbers}[2]{#1}
% \newcommand{\iflinenumbers}[2]{#2}
% \newcommand{\ifragged}[2]{#1}
\newcommand{\ifragged}[2]{#2}

\captionsetup[figure]{listformat=figList}

\usepackage[round]{natbib}
\usepackage{amsmath}
\title{The great codivergence bake off}

\author[1]{Tanner C.\ Myers}
\author[1]{Randy Klabacka}
\author[1]{Hailey Whitaker}
\author[1]{Tashitso Anamza}
\author[1]{Claire Tracy}
\author[1]{Jamie R.\ Oaks \thanks{Corresponding author: \href{mailto:joaks@auburn.edu}{\tt joaks@auburn.edu}}}
\affil[1]{Department of Biological Sciences \& Museum of Natural History, Auburn University, Auburn, Alabama 36849}

\date{\today}

\makeatletter
\let\msTitle\@title
\let\msAuthor\@author
\let\msDate\@date
\makeatother
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\ifragged{
\raggedright
}{}

\iflinenumbers{
\begin{linenumbers}
}{}

\textbf{Running head}: \uppercase{Codiv bake off}

{\let\newpage\relax\maketitle}

\newpage

\ifdoublespacing{
\doublespacing
}{}

\begin{abstract}
    Abstract here \ldots

    \vspace{6pt}
    \noindent\textbf{KEY WORDS:}
\end{abstract}

\newpage

\section{Introduction}

To understand the distribution of biodiversity, the field of historical
biogeography considers how past environmental changes have shaped diversity
within and among species.
Environmental processes that affect whole biological communities can cause
speciation or extinction across multiple species at the same time.
Biogeographers are often interested in testing for patterns of temporally
clustered speciation events as evidence that a shared process caused multiple
lineages to diverge.
For example, finding that a number of evolutionary lineages diverged
simultaneously can be difficult to explain via species-specific processes alone and
provides compelling evidence for a shared process of diversification, like the
fragmentation of an environment.

One way to test the predictions of such processes of diversification is to
infer the temporal pattern of divergences across multiple taxa, and determine
whether any subsets of the taxa shared the same divergence times.
Methods have been developed to test for such patterns of shared divergence
times across taxa using approximate Bayesian computation (ABC)
\citep{Hickerson2006,Huang2011,Oaks2014dpp}
and likelihood-based Bayesian inference
\citep{Oaks2018ecoevolity,Oaks2019codemog}.
Both approaches use Bayesian model choice to infer the temporal pattern of
divergences across multiple pairs of populations or species.

Several of the ABC methods were primarily developed for analyzing a small
number of unlinked genetic loci each comprising linked, nonrecombining sites.
These methods often struggle to detect multiple divergence times across pairs of
populations \citep{Oaks2012, Oaks2014reply} or have little information to
update \textit{a priori} expectations \citep{Oaks2014dpp}.
More fundamentally, the loss of information inherent to ABC approaches can
prevent them from discriminating among models
\citep{Robert2011,Marin2014,Green2015}.

By more efficiently using information from a large number of effectively
unlinked sites from across the genomes of sampled organisms, the
full-likelihood Bayesian methods enable more accurate and precise estimates of
shared divergneces
\citep{Oaks2018ecoevolity,Oaks2019codemog}.

This raises a question.
Given a dataset of one or a small number of loci of linked sites,
is it better to use an ABC approach that correctly models such loci, but
but suffers from low power and bias?
Or, is it better to use a full-likelihood approach that has more power,
but incorrectly assumes the sites in the data are unlinked?
\jrocomment{Need to flesh out this intro more.}

\section{Methods}

\subsection{Simulating data}

We simulated loci for five diploid individuals (i.e., 10 gene
copies) per population from five pairs of diverging populations.
Each pair shared a common ancestral population with a constant effective size
\epopsize[\rootpopindex]
that diverged at time
\comparisonetime{}
into two descendant populations with constant effective sizes
\epopsize[\descendantpopindex{1}]
and 
\epopsize[\descendantpopindex{2}].
Each locus we simulated comprised 500 linked characters with no intralocus
recombination.
We simulated datasets comprising 1, 2, 5, and 20 independent loci.
For each number of loci, we simulated datasets under three divergence
scenarios:
\begin{enumerate}
    \item All five pairs of populations diverged independently.
    \item All five pairs of populations diverged simultaneously.
    \item The number of divergence times (\nevents) and the assignment of the
        five pairs of populations to the divergence times (\etimesets) vary
        according to a Dirichlet-process model with a concentration parameter
        of 2.22543, which corresponds to a mean of 3 divergence times.
\end{enumerate}
For each of the 12 simulation conditions (4 numbers of loci $\times$ 3
divergence scenarios),
we simulated 1000 datasets.
Each unique divergence time was drawn from an exponential distribution,
$\comparisonetime \sim \dexponential{0.01}$.
We constrained the mutation rate (\murate{}) to 1 across the five pairs of
populations, such that divergence times and effective population sizes were
scaled by the mutation rate.

To simulate the datasets for analysis under a full-likelihood Bayesian model,
we used the \simcoevolity tool of the \ecoevolity software package.
\citep[Version 0.3.2, Commit c128046;][]{Oaks2018ecoevolity,Oaks2018paic}.
The mutation-scaled effective sizes of the descendant populations
($\epopsize\murate$) were drawn from a gamma distribution with a shape of 10
and mean of 0.0005,
$\epopsize{}\murate{} \sim \dgamma{10}{0.0005}$.
The effective size of the ancestral population relative to the mean of
its descendants was drawn from a gamma distribution with a shape of 40 and mean of 1,
$\rootrelativepopsize \sim \dgamma{40}{1.0}$.

To simulate datasets for analysis with
approximate-Bayesian computation (ABC),
we used \pymsbayes
\citep[Version 0.3.7, Commit 54f720df;][]{Oaks2014dpp},
which is a Python multiprocessing wrapper around \dppmsbayes
\citep{Oaks2014dpp}, which is a modification of
MTML-msBayes \citep{Huang2011}.
In this software, the distribution for effective population sizes is placed
on the parameter $4\epopsize{}\murate{}$, for which we used
a gamma distribution with a shape of 10 and a mean of 0.002.
This is equivalent to the distribution we used for $\epopsize{}\murate{}$ with
\ecoevolity.
When summarizing and plotting effective population size for \pymsbayes, we
divided all true and estimated values by four for direct comparability with
\ecoevolity.

Ideally, we would have applied both the full-likelihood and ABC methods
to the datasets simulated using \pymsbayes.
However, this was not possible, because \dppmsbayes does not store the full
simulated sequence data, only the population genetic summary statistics used by
the ABC algorithm.

\subsection{Full-likelihood analyses of simulated data}

We analyzed each dataset simulated by \simcoevolity with \ecoevolity
\citep[Version 0.3.2, Commit c128046;][]{Oaks2018ecoevolity,Oaks2018paic}
under a Dirichlet-process model with a concentration parameter of 2.22543.
For prior distributions on divergence times and effective population sizes, we
used the same distributions described above for the simulations.
We ran four MCMC chains for 30,000 generations sampling every 20 generations.
To approximate the posterior distribution of the model given each simulated
dataset, we used the last 1000 samples from each of the four chains.
To assess convergence and mixing of the MCMC chains, we calculated the
potential scale reduction factor \citep[PSRF; the square root of Equation 1.1
in][]{Brooks1998} and effective sample size \citep[ESS;][]{Gong2014} for all
continuous parameters and the log likelihood using
\pycoevolity
\citep[Version 0.2.11 Commit 85ea44b;][]{Oaks2018ecoevolity,PycoevolityRepoOnline}.

\subsubsection{ABC analyses of simulated data}

To analyze each dataset simulated by \pymsbayes, we used the ABC rejection
algorithm implemented in \dppmsbayes via \pymsbayes \citep[Version 0.3.7,
Commit 54f720df;][]{Oaks2014dpp},
under a Dirichlet-process model with a concentration parameter of 2.22543, and
prior distributions on divergence times and effective population sizes that
match the simulation distributions described above.
We simulated 1 million samples to approximate the joint prior distribution of
the model.
To approximate the posterior given each simulated dataset, we retained the 2000
prior samples with the smallest Euclidean distance for the following four
summary statistics standardized using the mean and standard deviation from the
prior sample
\citep{Oaks2014dpp}:
$\pi$ \cite{Tajima1983}, $\theta_W$ \cite{Watterson1975}, $\pi_{net}$
\cite{Takahata1985}, and $SD(\pi-\theta_W)$ \cite{Tajima1989}.

\subsection{Re-analyses of biological data}

To assess the effect of a using a full-likelihood approach, we reanalyzed two
datasets with previously published results of using the ABC approach to
estimate shared divergence times.

In their study into the factors affecting the divergence of flightless darkling beetles across the Aegean Archipelago,
Papadopoulou and Knowles \citeyear{PapadopoulouKnowles2015} used \dppmsbayes to estimate the relative divergence times of 13 pairs of populations that vary phenotypically and ecologically.
Their dataset included two Sanger loci --- one mitochondrial (Cox1; \textit{n} = 403) and one nuclear (Mp20; \textit{n} = 273)--- sequenced across hundreds of beetles from 60 localities on 11 islands of the Cycladic plateau in the Aegean archipelago \citep{PapadopoulouKnowles2015}.
\jrocomment{We should probably describe the dataset a bit; number of loci, etc.}

To test the prediction of clustered divergences driven by Pleistocene glacial
cycles Oaks et al.\ \citeyear{Oaks2012} used MTML-msBayes \citep{Huang2011} to
estimate the divergence times of 22 pairs of vertebrate populations from across the
Philippine Islands.
Oaks \citeyear{Oaks2014dpp} reanalyzed these same data using \dppmsbayes.
This dataset consists of one or two mitochondrial loci gathered from closely related species or populations within species distributed across contemporarily isolated islands that previous were contiguous.
\jrocomment{We should probably describe the dataset a bit; number of loci, etc.}

\jrocomment{Jamie needs to add analysis details here}.

\section{Results}

\subsection{Simulation results}

We found good convergence and mixing of the MCMC chains across the \ecoevolity
analyses of the 12,000 simulated dataset.
The potential scale reduction factor (PSRF) was less than 1.01 and the
effective sample size (ESS) was greater than 600 for all parameters and the log
likelihood across all simulations
(\figs S\labelcref{fig:psrfdivtime,fig:psrfrootpopsize,fig:psrflnl,fig:essdivtime,fig:essrootpopsize,fig:esslnl}).

Across all four numbers of loci and three divergence scenarios, \ecoevolity
performed better at estimating the divergence model (\ie, the number of
divergence times and the assignment of the five pairs of populations to those
divergences) than \dppmsbayes, and the differences were significant according
to a Mann-Whitney U test \citep[\fig~\ref{fig:modelerror};][]{MannWhitney1947}.
The better performance of \ecoevolity is small when there is only one 500-base
locus, but the performance difference increases as the number of loci increase.
This is largely due to the fact that the ABC method (\dppmsbayes) does not gain
much information from additional loci, whereas the full-likelihood method
(\ecoevolity) does (\fig~\ref{fig:modelerror}).
The same pattern is true when looking at estimation error of divergence times
(\figs \ref{fig:divtimeerror}, \ref{fig:divtimescatterdp}, \& S\ref{fig:divtimescatter}),
the number of divergence times
(\fig S\ref{fig:nevents}),
and the effective population sizes 
(\figs S\labelcref{fig:ancpopsizeerror,fig:leafpopsizescatter,fig:rootpopsizescatter}).

\begin{figure}[htbp]
    \begin{center}
        \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{../images/from-project-repo/plots/tex-plot-grids/grid-model-distance-cropped.pdf}
        \caption{
            \Ecoevolity estimates the partitioning of population pairs to
            divergence times (the divergence model) more accurately and
            precisely than \dppmsbayes across all simulated numbers of loci
            (columns) and whether the divergence model was drawn from a
            Dirichlet process (top row), or constrained so that all pairs of
            populations diverged independently (middle row) or simultaneously
            (bottom row).
            Divergence model estimation error was quantified for each
            simulation replicate using the mean partition distance
            \citep{Regnier1983,Gusfield2002} between the true divergence model
            and the posterior samples of divergence models.
            Partition distances was calculated using the Hungarian (or
            Kuhn-Munkres) algorithm
            \citep{Kuhn1955,Munkres1957}
            as implemented in \pycoevolity
            \citep[Version 0.2.11 Commit 85ea44b;][]{Oaks2018ecoevolity,PycoevolityRepoOnline},
            which relies on the Munkres Python package
            \citep[Version 1.1.4;][]{Clapper2020}.
            For each violin plot, the sample size (number of simulation
            replicates) is shown beneath and the mean is represented by the
            white dot and shown above.
            The box shows the upper and lower quartiles and the brackets
            show the 2.5th and 97.5th percentiles.
            At the top center of each plot, the results of a Mann-Whitney U
            test \citep{MannWhitney1947} are shown comparing mean model errors
            between \ecoevolity and \dppmsbayes;
            the probability of superiority \citep[PoS, the probability that
            \ecoevolity has a lower mean model error than \dppmsbayes for a
            random simulation replicate drawn from
            each;][]{WolfeHogg1971,Grissom1994} and p-value is provided.
        }
        \label{fig:modelerror}
    \end{center}
\end{figure}

\begin{figure}[htbp]
    \begin{center}
        \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{../images/from-project-repo/plots/tex-plot-grids/grid-div-time-error-cropped.pdf}
        \caption{
            \Ecoevolity better estimates divergence times than \dppmsbayes
            across all simulated numbers of loci (columns) and whether the
            divergence model was drawn from a Dirichlet process (top row), or
            constrained so that all pairs of populations diverged independently
            (middle row) or simultaneously (bottom row).
            Divergence time error was quantified for each simulation replicate
            as the sum of the difference between the true and posterior mean
            divergence time across the five pairs of populations.
            For each violin plot, the sample size (number of simulation
            replicates) is shown beneath and the mean is represented by the
            white dot and shown above.
            The box shows the upper and lower quartiles and the brackets
            show the 2.5th and 97.5th percentiles.
            At the top center of each plot, the results of a Mann-Whitney U
            test \citep{MannWhitney1947} are shown comparing mean model errors
            between \ecoevolity and \dppmsbayes;
            the probability of superiority \citep[PoS, the probability that
            \ecoevolity has a lower mean model error than \dppmsbayes for a
            random simulation replicate drawn from
            each;][]{WolfeHogg1971,Grissom1994} and p-value is provided.
        }
        \label{fig:divtimeerror}
    \end{center}
\end{figure}

\begin{figure}[htbp]
    \begin{center}
        \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{../images/from-project-repo/plots/free-div-time-scatter-cropped.pdf}
        \caption{
            \Ecoevolity (top row) estimates divergence times more accurately
            and precisely than \dppmsbayes (bottom row) across all simulated
            numbers of loci (columns) for five pairs of populations simulated
            to have diverged according to a Dirichlet process.
            Each plotted circle and associated error bars represent the posterior mean
            and 95\% credible interval.
            \accuracyscatterplotannotations{\comparisonetime{}}
        }
        \label{fig:divtimescatterdp}
    \end{center}
\end{figure}

% We can remove this page break later, just adding it now to make it easier to
% delineate empirical results in the PDF
\newpage

\subsection{Empirical results}

\jrocomment{For now I am just dropping plots below.}

\begin{figure}[htbp]
    \begin{center}
        \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{../images/from-project-repo/empirical-plots/mediterranean-all-pairs/pyco-sumevents-mediterranean-all-pairs-pycoevolity-nevents-cropped.pdf}
        \caption{
            \Ecoevolity results for all Mediterranean pairs.
        }
        \label{fig:neventsmediall}
    \end{center}
\end{figure}

\begin{figure}[htbp]
    \begin{center}
        \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{../images/from-project-repo/empirical-plots/mediterranean-soil-pairs/pyco-sumevents-mediterranean-soil-pairs-pycoevolity-nevents-cropped.pdf}
        \caption{
            \Ecoevolity results for soil Mediterranean pairs.
        }
        \label{fig:neventsmedisoil}
    \end{center}
\end{figure}

\begin{figure}[htbp]
    \begin{center}
        \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{../images/from-project-repo/empirical-plots/philippines-all-pairs/pyco-sumevents-philippines-all-pairs-pycoevolity-nevents-cropped.pdf}
        \caption{
            \Ecoevolity results for all Philippines pairs.
        }
        \label{fig:neventsphilall}
    \end{center}
\end{figure}

\begin{figure}[htbp]
    \begin{center}
        \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{../images/from-project-repo/empirical-plots/philippines-negros-panay-pairs/pyco-sumevents-philippines-negros-panay-pairs-pycoevolity-nevents-cropped.pdf}
        \caption{
            \Ecoevolity results for Negros-Panay pairs.
        }
        \label{fig:neventsphilnp}
    \end{center}
\end{figure}

% We can remove this page break later, just adding it now to make it easier to
% delineate empirical results in the PDF
\newpage

\section{Discussion}

As mentioned in the methods,
it would have been ideal to apply both the full-likelihood and ABC methods
to the data simulated via \dppmsbayes.
Because \dppmsbayes only stores summary statistics from the genetic data, this
was not possible.
However, it is unlikely that differences between the simulators implemented in
\dppmsbayes and \ecoevolity could explain the differences in performance
between full-likelihood and ABC inference methods.
Both simulators have been independently validated by suites of tests in the
\pymsbayes and \ecoevolity distributions.
Also, the assumption that the simulators are producing nearly identically
distributed data (barring minor differences in parameterizations between the
implementations) is supported by our finding that both inference methods
perform very similarly when the size of the simulated data sets are small
(i.e., one 500-site locus) and diverge in performance as we increased the size
of the simulated data sets (\figs \ref{fig:modelerror} \&
\ref{fig:divtimeerror}).

\section{Acknowledgments}

This work was supported by funding provided to JRO from the National Science
Foundation (NSF grant number DEB 1656004).
The computational work was made possible by the Auburn University (AU) Hopper
Cluster supported by the AU Office of Information Technology and a grant of
high-performance computing resources and technical support from the Alabama
Supercomputer Authority.
This paper is contribution number \highLight{XXX} of the Auburn University
Museum of Natural History.

%TC:ignore

\bibliographystyle{evolution}
\bibliography{references}

\iflinenumbers{
\end{linenumbers}
}{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% SUPPORTING INFO %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\setcounter{figure}{0}
\setcounter{table}{0}
\setcounter{page}{1}
\setcounter{section}{0}

\captionsetup[figure]{name=Figure S, labelformat=noSpace, listformat=sFigList}
\captionsetup[table]{name=Table S, labelformat=noSpace}

\singlespacing

\section*{Supporting Information}
\hangindent=1cm
\noindent Title: \msTitle

\bigskip
{\noindent Authors: \msAuthor}

\newpage
\singlespacing

\input{si-body.tex}
\clearpage

\input{si-tables.tex}
\clearpage

\input{si-figures.tex}

%TC:endignore

\end{document}